---
title: 'hw2: the beta family of pdf''s'
author: "Keoishia Grigsby"
output:
  html_notebook: default
  pdf_document: default
---

# Part 1
Write your own version of `dbeta()` using the beta pdf formula given above. Call your function `mydbeta()`.  Use three arguments: `x`, `shape1` for $\alpha$, and `shape2` for $\beta$.  Don't bother with `ncp` or `log`. 
```{r}
mydbeta <- function(x, shape1, shape2){
  #dbeta(x, shape1, shape2) 
  ((1/beta(shape1,shape2)) * (x**(shape1 - 1)) * ((1 - x)**(shape2 - 1)))
}
```

Test `mydbeta()`  by comparing its output to `dbeta()` output.  Choose 5 sets of values for $x, \alpha$ and $\beta$ and make a data.frame to summarize your results.  
Plot the pdf's of your five test cases together on the same plot.    

```{r}
x <- runif(5, 0, 1)
shape1 <- runif(5, 0, 100)
shape2 <- runif(5, 0, 100)


mydata <- cbind.data.frame(x, shape1, shape2)
mydata['mydbeta'] <- mydbeta(mydata[, 1], mydata[, 2], mydata[, 3])
mydata['dbeta'] <- dbeta(mydata[, 1], mydata[, 2], mydata[, 3])
mydata
```


# Part 2
Use your function `mydbeta()` for these questions. 
Experiment with different values of $\alpha$ and $\beta$ and plot the resulting beta pdf's to answer the following questions. The parameters $\alpha$ and $\beta$ have to be greater than zero. Otherwise there's no restriction.  

(@) Find values for $\alpha$ and $\beta$ which produce a pdf which is uniform (flat) on [0,1].    _______________________ Make a plot of this beta density. 

* The values of alpha and beta that make a uniform distribution are (1, 1)
```{r}
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = 1, shape2 = 1) # flat
data1 <- cbind.data.frame(xvec, yvec)
data1
plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(0,2))
title('Beta distribution with uniform distribution')
```


(@) What relationship between $\alpha$ and $\beta$ produces a symmetric pdf? ____________________________ Make a plot with three examples of such a pdf. 
* When alpha and beta are equal their pdf is symmertric 
```{r}
# first line
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = 50, shape2 = 50)
#second line
xvec2 <- seq(0, 1, by = .01)
yvec2 <- mydbeta(xvec, shape1 = 5, shape2 = 5)
#third line
xvec3 <- seq(0, 1, by = .01)
yvec3 <- mydbeta(xvec, shape1 = 10, shape2 = 10)


mydata2 <- cbind.data.frame(xvec, yvec, yvec2, xvec3, yvec3)
mydata2
plot(xvec, yvec, type='l', col = '6', lwd = 'k', xlim=c(0,1), ylim = c(-1,10))
lines(xvec2, yvec2, col = 9)
lines(xvec3, yvec3, col = 27)
title('Beta distribution with symmertric distribution')
```


(@) What relationship between $\alpha$ and $\beta$ produces a pdf which is skewed to the right? _________________________ Make a plot with three examples of such a pdf.  

* when beta is greater than alpha the graph is skewed to the right
```{r}
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = 30, shape2 = 70)

xvec2 <- seq(0, 1, by = .01)
yvec2 <- mydbeta(xvec, shape1 = 3, shape2 = 7)

xvec3 <- seq(0, 1, by = .01)
yvec3 <- mydbeta(xvec, shape1 = 3, shape2 = 17)

mydata3 <- cbind.data.frame(xvec, yvec, xvec2, yvec2, xvec3, yvec3)
mydata3

plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(-1,10))
lines(xvec2, yvec2, col = 9)
lines(xvec3, yvec3, col = 27)
title('Beta Distribution with Right Skwedness')
```

(@) What values for $\alpha$ and $\beta$ produce a triangular pdf?____________________________  Make a plot with two examples of such a pdf. 
* for this one I could ony guess when x only has 3 values or if the x values goi by .05
```{r}
# first line
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = 1.000001, shape2 = 2.000001)

#second line
xvec2 <- seq(0, 1, by = .01)
yvec2 <- mydbeta(xvec2, shape1 = 2.000001, shape2 = 1.000001)

#saved df of data
mydata4 <- cbind.data.frame(xvec, yvec, xvec2, yvec2)
mydata4

#plots of lines
plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(0,2))
lines(xvec2, yvec2, col = 3)
title('Beta distribution with triangular peak')
```


(@) We say **spike** to describe a function where most values are close to zero except for a tall skinny peak.  How should I choose $\alpha$ and $\beta$ to get a beta pdf with a spike at x = .5, and close to 0 everywhere else?  ____________________________  Make a plot with two examples of such a pdf. 

* a very large alpha and beta should be schosen if we want to see a spike
```{r}
xvec <- seq(0, 1, by = .00001)
yvec <- mydbeta(xvec, shape1 =500 , shape2 = 500 )

xvec2 <- seq(0, 1, by = .00001)
yvec2 <- mydbeta(xvec, shape1 = 250, shape2 = 250)

mydata5 <- cbind.data.frame(xvec, yvec, xvec2, yvec2)
mydata5

plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(-1,10))
lines(xvec2, yvec2, col = 9)
title('Beta Distribution with Spike at x =0.5')
```

(@) Suppose I want spikes near x = 0 and x = 1 with density close to 0 in between? ____________________________  Make a plot with three examples of such a pdf.
```{r}
#first line
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = .005 , shape2 =.005)
#second line
xvec2 <- seq(0, 1, by = .001)
yvec2 <- mydbeta(xvec2, shape1 = .08, shape2 = .08)
#third linbe
xvec3 <- seq(0, 1, by = .001)
yvec3 <- mydbeta(xvec3, shape1 = .001, shape2 = .001)
#plots

plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(-1,10))

lines(xvec2, yvec2,  col = 4)
lines(xvec3, yvec3,  col = 39)

#data
mydata5 <- cbind.data.frame(xvec, yvec, xvec2, yvec2)
mydata5

title('Beta Distribution with Spike at x = 0 and x =5 ')
```

(@) What values for $\alpha$ and $\beta$ produce a pdf which is bell-shaped?  ____________________________ It should look close to normal.   Make a plot with at least one example of such a pdf. 
* (3, 3) and (2, 2) for alpha and beta produce a bell-shape
```{r}
#bell-shaped lkine
xvec <- seq(0, 1, by = .01)
yvec <- mydbeta(xvec, shape1 = 3, shape2 = 3)


plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(0,2))

mydata6 <- cbind.data.frame(xvec, yvec)
mydata6

title('Beta Distribution with "Bell-Shape"')
```


(@) Find values for $\alpha$ and $\beta$  which produce a pdf like a concave bowl. (Concave means curved like a frown.  Curved like a smile is convex.)  ____________________________  Make a plot with three examples of such a pdf. 
* when alpha and beta are equal and very small the pdf shape is convex
```{r}
xvec <- seq(0, 1, by = .001)
yvec <- mydbeta(xvec, shape1 = .001, shape2 = .001)

xvec2 <- seq(0, 1, by = .001)
yvec2 <- mydbeta(xvec, shape1 = .01, shape2 = .01)
plot(xvec, yvec, type='l', col = 6, lwd = 'k', xlim=c(0,1), ylim = c(0,3))

lines(xvec2, yvec2,  col = 4)


title('Beta Distribution with convex')

```